---
title: "WUP_04"
author: "Daniel"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSS = read.csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->



# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

I used chatgpt to get R code for plotting

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
#Load necessary libraries (if not already loaded)
#install.packages('ggplot2') # Install ggplot2 if needed
library(ggplot2)

#create a box plot
ggplot(data = GSS, aes(x =  DEGREE, y = WORDSUM)) +
  geom_boxplot() +
  xlab("DEGREE") +
  ylab("WORDSUM") +
  ggtitle("Boxplot of Response Varible by Category")

```

With the plot we can see that the average score varies, but will go up with a degree. Those with a higher degree, on average, have a higher WORDSUM score.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data = GSS)
```

The closest in mean and median would be BACHALOR and JUNIOR COLLEGE, but even then there is a difference of .4 between them. This would indicate that even though there is difference seeing as the entirety of the means and medians all vary, showing a correlation between WORDSUM score and DEGREE.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data = GSS)
summary.aov(model1)
```
```{r}
plot(TukeyHSD(model1))
print(TukeyHSD(model1))
```

The graph will have our given p-value of 0.000236, which is lower than 0.05, which means we reject the null hypothesis. There is a difference in WORDSUM score and DEGREE. The differences would be Jr college to Bachelors, Jr College to High School, High School to Graduate, and LIHS to Graduate.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

During our research and information gathered. The question we had was, is there a difference between the average score on the WORDSUM test and level of schooling DEGREE? With our findings we can say there is a difference. During our TuckyHSD test, he receive a p-value of 0.000236, which is less than 0.05. IN turn we reject the null hypothesis of no difference and conclude there is a difference. Further in we find that the differences most noticable would be Jr College to Bachelors, Jr College to High School, High School to Graduate, and LIHS to Graduate. 
